Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-11 Manuscript/manuscript.tex
A small commit fix

In this work we show that the cation binding behavior
in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
improved by implicitly including electronic polarizability
in the polar region of lipid molecules. The electronic
polarizability is included by using the electronic
continuum correction (ECC)~\cite{leontyev11}, which
has been previously shown to improve the behaviour of
MD simulations of ions in bulk water~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
As a starting point we use the parameters from Lipid14 model \cite{dickson14},
which gave the best cation binding behaviour in the previous study \cite{catte16}.
The developed ECC-lipid parameters reproduce the experimentally
measurable structural parameters of POPC lipid bilayer with the 
accuracy comparable to the other state of the art lipid models.

@@

The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
From recent theoretical work with long enough simulations this time can be roughly estimated
in the order of $1$--$10\,\mu\mathrm{s}$~\cite{Javanainen2017}. 
This is in contrast to our model, ECC-lipids, 
which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
\todo{evaluate this number, mean residence time, accurately based on the contacts data.}
i.e. at least two orders of magnitude lower than previous estimates.
Such a finding changes the point of view of calcium binding from
very tight long-term stable binding with rare exchanges to 
a relatively frequent exchange of cations 
in equilibrium between membrane and solvent.

>>>>

In this work we show that the cation binding behavior
in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
improved by implicitly including electronic polarizability
in the polar region of lipid molecules. The electronic
polarizability is included by using the electronic
continuum correction (ECC)~\cite{leontyev11}, which
has been previously shown to improve the behaviour of
MD simulations of ions in bulk water~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
As a starting point we use the parameters from Lipid14 model \cite{dickson14},
which gave the best cation binding behaviour in the previous study \cite{catte16}.
The developed ECC-lipid parameters reproduce the experimentally
measurable structural parameters of POPC lipid bilayer with the 
accuracy comparable to the other state of the art lipid models.

@@

The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
From recent theoretical work with long enough simulations this time can be roughly estimated
in the order of $1$--$10\,\mu\mathrm{s}$~\cite{javanainen17}. 
This is in contrast to our model, ECC-lipids, 
which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
\todo{evaluate this number, mean residence time, accurately based on the contacts data.}
i.e. at least two orders of magnitude lower than previous estimates.
Such a finding changes the point of view of calcium binding from
very tight long-term stable binding with rare exchanges to 
a relatively frequent exchange of cations 
in equilibrium between membrane and solvent.
